---
title: "Process NCI data"
author: "Clemens Hug"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(vroom)
library(fst)
library(here)

dir_data <- here("data")
dir.create(dir_data, showWarnings = FALSE)
dir_res <- here("nci")
dir.create(dir_res, showWarnings = FALSE)

synapser::synLogin()
syn <- synExtra::synDownloader(dir_data)
```

## Download

```{r download, message=FALSE}
options(timeout = 300)

nci_60_doseresp_path <- file.path(dir_data, "nci_60_doseresp.zip")
if (!file.exists(nci_60_doseresp_path))
  download.file(
    "https://wiki.nci.nih.gov/download/attachments/147193864/DOSERESP.zip?version=1&modificationDate=1622830743000&api=v2",
    nci_60_doseresp_path
  )

nci_almanac_path <- file.path(dir_data, "nci_almanac.zip")
if (!file.exists(nci_almanac_path))
  download.file(
    "https://wiki.nci.nih.gov/download/attachments/338237347/ComboDrugGrowth_Nov2017.zip?version=1&modificationDate=1510057275000&api=v2",
    nci_almanac_path
  )

nci_almanac_drug_meta_path <- file.path(dir_data, "nci_almanac_drug_meta.csv")
if (!file.exists(nci_almanac_drug_meta_path))
  download.file(
    "https://wiki.nci.nih.gov/download/attachments/338237347/ComboCompoundNames_small.txt?version=1&modificationDate=1493822467000&api=v2",
    nci_almanac_drug_meta_path
  )

unzip(nci_60_doseresp_path)
unzip(nci_almanac_path)
```

```{r read}
lspci_drug_names <- syn("syn25173503") %>%
  read_fst(as.data.table = TRUE) %>%
  mutate(across(name, str_to_lower))

lspci_clinical_info <- syn("syn25173555") %>%
  read_fst(as.data.table = TRUE)

lspci_compound_dictionary <- syn("syn25173500") %>%
  read_fst(as.data.table = TRUE)

nci_60_doseresp_raw <- file.path(dir_data, "DOSERESP.csv") %>%
  vroom()

nci_almanac_raw <- file.path(dir_data, "ComboDrugGrowth_Nov2017.csv") %>%
  vroom()

nci_almanac_drug_meta_raw <- read_tsv(nci_almanac_drug_meta_path, col_names = c("id", "name"))
```

```{r map_drugs}
nci_almanac_drug_meta <- nci_almanac_drug_meta_raw %>%
  mutate(across(name, str_to_lower)) %>%
  # Rename manually mapped drugs
  mutate(
    name = recode(name, `2-fluoro ara-a` = "fludarabine", `4'-epiadriamycin` = "doxorubicin"),
    split = FALSE
  ) %>%
  bind_rows(
    mutate(., name = str_split(name, fixed(" ")) %>% map_chr(1), split = TRUE)
  ) %>%
  left_join(lspci_drug_names, by = "name") %>%
  drop_na() %>%
  distinct(id, lspci_id, source, priority, split) %>%
  left_join(
    lspci_clinical_info %>%
      mutate(max_phase = coalesce(max_phase, 0)) %>%
      distinct(lspci_id, max_phase, indication_class) %>%
      group_by(lspci_id) %>%
      filter(max_phase == max(max_phase)) %>%
      ungroup()
  ) %>%
  mutate(max_phase = coalesce(max_phase, 0)) %>%
  # If there are matches without splitting names, keep only those
  # If there are still multiple matches, keep one in highest clinical phase
  group_by(id) %>%
  filter(if (any(!split)) !split else TRUE, max_phase == max(max_phase, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct(id, lspci_id, max_phase) %>%
  left_join(
    lspci_compound_dictionary %>%
      select(lspci_id, pref_name)
  )

write_csv(
  nci_almanac_drug_meta,
  file.path(dir_res, "nci_almanac_drug_meta.csv")
)
```

PercentGrowth    NUMBER           percent growth of drug combination 
                                   against cell line, using time zero in 
                                   calculation
PercentGrowthNoTZ NUMBER          percent growth without time zero
TestValue         NUMBER          test mean optical density
ControlValue      NUMBER          vehicle control mean optical density
TZValue           NUMBER          time zero mean optical density
ExpectedGrowth    NUMBER          expected percent growth for combination, 
                                   based on values for individual drugs
Score             NUMBER(4)       score, for combination records


```{r wrangle_almanac}
cell_line_meta <- nci_almanac_raw %>%
  distinct(
    panel_id = PANELNBR, cell_id = CELLNBR, panel_name = PANEL, cell_name = CELLNAME
  )

write_csv(
  cell_line_meta,
  file.path(dir_res, "nci_almanac_cell_line_meta.csv")
)

nci_almanac <- nci_almanac_raw %>%
  select(
    # All concentrations are in M, leaving column out
    panel_id = PANELNBR,
    cell_id = CELLNBR,
    drug_id_1 = NSC1,
    drug_concentration_1 = CONC1,
    drug_concentration_index_1 = CONCINDEX1,
    drug_id_2 = NSC2,
    drug_concentration_2 = CONC2,
    drug_concentration_index_2 = CONCINDEX2,
    PERCENTGROWTH,
    PERCENTGROWTHNOTZ,
    TESTVALUE,
    CONTROLVALUE,
    TZVALUE,
    EXPECTEDGROWTH,
    SCORE
  ) %>%
  left_join(
    select(nci_almanac_drug_meta, drug_id_1 = id, lspci_id_1 = lspci_id)
  ) %>%
  left_join(
    select(nci_almanac_drug_meta, drug_id_2 = id, lspci_id_2 = lspci_id)
  )

fwrite(
  nci_almanac,
  file.path(dir_res, "nci_almanac_growth_data.csv.gz")
)
```


```{r store_synapse}
activity <- synapser::Activity(
  "Wrangle NCI growth data",
  used = c(
    "syn25173503", "syn25173555", "syn25173500",
    "https://wiki.nci.nih.gov/display/NCIDTPdata/DTP+NCI+Bulk+Data+for+Download"
  ),
  executed = "https://github.com/clemenshug/brca_synergy/blob/main/R/nci_data.Rmd"
)

syn_project <- "syn25882516"
syn_nci <- synExtra::synMkdir(syn_project, "data", "nci", .recursive = TRUE)

c(
  file.path(dir_res, "nci_almanac_drug_meta.csv"),
  file.path(dir_res, "nci_almanac_cell_line_meta.csv"),
  file.path(dir_res, "nci_almanac_growth_data.csv.gz")
) %>%
  synExtra::synStoreMany(parentId = syn_nci, forceVersion = FALSE, activity = activity)

```
