---
title: "Plot NCI data"
author: "Clemens Hug"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(here)
library(ggbeeswarm)
library(patchwork)

theme_set(theme_minimal())

dir_data <- here("data")
dir.create(dir_data, showWarnings = FALSE)
dir_res <- here("nci")
dir.create(dir_res, showWarnings = FALSE)

synapser::synLogin()
syn <- synExtra::synDownloader(dir_data, ifcollision = "overwrite.local")
```

## Load data from Synapse

```{r load}
growth_data_raw <- syn("syn25882596") %>%
  fread()
meta_cell_line <- syn("syn25882595") %>%
  fread()
meta_drugs <- syn("syn25882594") %>%
  fread() %>%
  # There are duplicates where multiple drug ids in the dataset map to the same lspci_id
  # Getting rid of duplicates for this analysis here
  distinct(lspci_id, max_phase, pref_name)
```

## Format data for plotting

PercentGrowth    NUMBER           percent growth of drug combination
against cell line, using time zero in
calculation
PercentGrowthNoTZ NUMBER          percent growth without time zero
TestValue         NUMBER          test mean optical density
ControlValue      NUMBER          vehicle control mean optical density
TZValue           NUMBER          time zero mean optical density
ExpectedGrowth    NUMBER          expected percent growth for combination,
based on values for individual drugs
Score             NUMBER(4)       score, for combination records

```{r qc}
StatPercentileX <- ggproto("StatPercentileX", Stat,
  compute_group = function(data, scales, probs) {
    percentiles <- quantile(data$x, probs=probs)
    data.frame(xintercept=percentiles, x=percentiles)
    },
  required_aes = c("x")
)

stat_percentile_x <- function(mapping = NULL, data = NULL, geom = "vline",
                              position = "identity", na.rm = FALSE,
                              show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    stat = StatPercentileX, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

diagnostic_histogram <- function(df, x) {
  x_sym <- enquo(x)
  df %>%
    ggplot(aes(!!x_sym)) +
      stat_percentile_x(geom = "linerange", ymin = -0.05, ymax = 0, probs = seq(.1, 0.9, by = 0.1), color = "red") +
      # geom_linerange(
      #   aes(q, ymin = -0.05, ymax = 0),
      #   data = ~quantile(.x[[rlang::as_name(x_sym)]], seq(.1, 0.9, by = 0.1), na.rm = TRUE) %>%
      #     enframe("quantile", "q"),
      #   color = "red"
      # ) +
      geom_histogram(aes(y = stat(ndensity)), bins = 50)
      # geom_density(color = "red")
}

plt <- growth_data_raw %>%
  diagnostic_histogram(CONTROLVALUE) +
  scale_x_log10() +
  facet_wrap(vars(study_location))
ggsave(file.path(dir_res, "nci_almanac_CONTROLVALUE_histogram.pdf"), plt, width = 9, height = 3)

plt <- growth_data_raw %>%
  diagnostic_histogram(TESTVALUE) +
  scale_x_log10() +
  facet_wrap(vars(study_location))
ggsave(file.path(dir_res, "nci_almanac_TESTVALUE_histogram.pdf"), plt, width = 9, height = 3)

plt <- growth_data_raw %>%
  diagnostic_histogram(TZVALUE) +
  scale_x_log10() +
  facet_wrap(vars(study_location))
ggsave(file.path(dir_res, "nci_almanac_TZVALUE_histogram.pdf"), plt, width = 9, height = 3)

plt <- growth_data_raw %>%
  count(study_location, measurement_type = if_else(is.na(lspci_id_2), "single_agent", "combination")) %>%
  ggplot(aes(study_location, n, fill = measurement_type)) +
    geom_col(position = "dodge")
ggsave(file.path(dir_res, "nci_almanac_study_location_counts.pdf"), plt, width = 5, height = 3)

x <- growth_data_raw %>%
  as.data.table() %>% {
    .[, if (.N > 1) .SD, by = c("panel_id", "cell_id", "lspci_id_1", "lspci_id_2", "drug_concentration_1", "drug_concentration_2")]
  }
# We have often 2 replicates per drug combo, concentration combo
# Have to average them

y <- x[, if(length(unique(study_location)) > 1) .SD, by = c("panel_id", "cell_id", "lspci_id_1", "lspci_id_2", "drug_concentration_1", "drug_concentration_2")]
# Often multiple study locations per drug combo, concentration combo
# Have to be careful aggregating data because raw values are influenced by assay type
# Can't do simple average of raw values

growth_data_raw %>%
  group_by(
    lspci_id_1, measurement_type = if_else(is.na(lspci_id_2), "single_agent", "combination")
  ) %>%
  summarize(n_locations = length(unique(study_location)), measurement_types = .groups = "drop")
```



```{r gr}
compute_metrics <- function(data) {
  mutate(
    data,
    fraction_killing = pmax(0, 1 - (TESTVALUE / CONTROLVALUE)),
    log_kills = pmax(0, log10(CONTROLVALUE / TESTVALUE)),
    fraction_killing = 1 - (TESTVALUE / CONTROLVALUE),
    log_kills = log10(CONTROLVALUE / TESTVALUE),
    gr = log2(TESTVALUE / TZVALUE) / log2(CONTROLVALUE / TZVALUE),
    gr_bound = 2**gr - 1
  )
}

compute_bliss <- function(data) {
  mutate(
    data,
    # Cap the fraction killing between 0 and 1 for bliss model
    fraction_killing_bliss_expected = fraction_killing_1 + fraction_killing_2 - fraction_killing_1 * fraction_killing_2,
    excess_bliss = log10(1 - fraction_killing_bliss_expected) - log10(1 - fraction_killing),
    log_kills_bliss_expected = log10(1 / (1 - fraction_killing_bliss_expected)),
    gr_bliss_expected = 
  )
}

# harmonic_mean <- function(x) {
#   1 / mean(1 / x)
# }

growth_data_all <- growth_data_raw %>%
  # # Remove data with extremely low control or time zero counts
  # # Actually, even the low values seem to be fine. They result from
  # # a different protocol being used by one of the study centers
  # # So we only have to filter for negative values
  # filter(
  #   CONTROLVALUE > 1e3,
  #   TZVALUE > 1e3
  # ) %>%
  filter(TESTVALUE > 0, !study_location %chin% c("1A")) %>%
  compute_metrics()

growth_data_filtered <- growth_data_all %>%
  group_by(panel_id, cell_id, lspci_id_1, lspci_id_2, drug_concentration_1, drug_concentration_2) %>%
  filter(
    study_location %chin% names(which.max(table(study_location)))
  ) %>%
  filter(
    n() > 1
  ) %>%
  filter(
    t.test(gr, mu = 1, alternative = "less")$p.value < 0.05
  )

growth_data <- growth_data_filtered %>%
  summarize(
    across(
      where(is.numeric), mean
    ), .groups = "drop"
  )
  # setDT() %>% {
  #   .[
  #     study_location == names(which.max(table(study_location))),
  #     keyby = c("panel_id", "cell_id", "lspci_id_1", "lspci_id_2", "drug_concentration_1", "drug_concentration_2")
  #   ][
  #     ,
  #     lapply(.SD, mean),
  #     .SDcols = c("fraction_killing", "log_kills", "gr", "gr_bound"),
  #     keyby = c("panel_id", "cell_id", "lspci_id_1", "lspci_id_2", "drug_concentration_1", "drug_concentration_2")
  #   ]
  # }
  # as.data.table() %>% {
  #   .[
  #     ,
  #     lapply(.SD, mean),
  #     by = c("panel_id", "cell_id", "lspci_id_1", "lspci_id_2", "drug_concentration_1", "drug_concentration_2"),
  #     .SDcols = c("fraction_killing", "log_kills", "gr", "gr_bound")
  #   ]
  # }

growth_data_single_dose_filtered <- growth_data_filtered %>%
  ungroup() %>%
  filter(is.na(lspci_id_2)) %>%
  rename(
    lspci_id = lspci_id_1,
    drug_concentration = drug_concentration_1
  )

growth_data_single_dose_agg <- growth_data %>%
  filter(is.na(lspci_id_2)) %>%
  rename(
    lspci_id = lspci_id_1,
    drug_concentration = drug_concentration_1
  )
  
growth_data_bliss_raw <- growth_data %>%
  filter(!is.na(lspci_id_2)) %>%
  left_join(
    growth_data_single_dose_agg %>%
      select(
        panel_id, cell_id,
        lspci_id_1 = lspci_id, drug_concentration_1 = drug_concentration,
        fraction_killing_1 = fraction_killing,
        gr_1 = gr,
        log_kills_1 = log_kills
      )
  ) %>%
  left_join(
    growth_data_single_dose_agg %>%
      select(
        panel_id, cell_id,
        lspci_id_2 = lspci_id, drug_concentration_2 = drug_concentration,
        fraction_killing_2 = fraction_killing,
        gr_2 = gr,
        log_kills_2 = log_kills
      )
  ) %>%
  # Some drug combos don't have matching single agent data at the same concentration
  # Dropping them
  drop_na(fraction_killing_1, fraction_killing_2) %>%
  compute_bliss()

### Calculate stats of how often some compounds are synergistic

plt <- growth_data_all %>%
  diagnostic_histogram(gr) +
    lims(x = c(-2, NA)) +
    facet_wrap(~study_location)

ggsave(
  file.path(dir_res, "nci_almanac_gr_histogram.pdf"), plt,
  width = 6, height = 2
)
```

```{r plot_bliss}
growth_data_bliss_nested <- growth_data_bliss_raw %>%
  group_nest(
    panel_id, cell_id, lspci_id_1, lspci_id_2
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  left_join(
    distinct(meta_drugs, lspci_id_1 = lspci_id, pref_name_1 = pref_name)
  ) %>%
  left_join(
    distinct(meta_drugs, lspci_id_2 = lspci_id, pref_name_2 = pref_name)
  )

adam_cmap <- c(
  "#FFDA64", "#FFC952", "#FAB534", "#FCAB24", "#F79820", "#F48221", "#F4751C",
  "#EB6820", "#D95330", "#D04445", "#B93B51", "#953565", "#78287E", "#522B6E", "#22152F"
)

plot_gr <- function(panel_id, cell_id, lspci_ids, y = gr, highlight_conc = FALSE) {
  y_sym <- enquo(y)
  data_agg <- growth_data_single_dose_agg %>%
    filter(panel_id == !!panel_id, cell_id == !!cell_id, lspci_id %in% lspci_ids) %>%
    left_join(meta_drugs)
  data_all <- growth_data_single_dose_filtered %>%
    filter(panel_id == !!panel_id, cell_id == !!cell_id, lspci_id %in% lspci_ids) %>%
    left_join(meta_drugs)
  plt <- ggplot(data_agg, aes(drug_concentration, !!y_sym)) +
    geom_line(color = "gray") +
    geom_beeswarm(aes(color = study_location), data = data_all, groupOnX = TRUE, alpha = 0.7) +
    scale_x_log10() +
    labs(x = "Concentration") +
    theme_minimal()
  if (length(lspci_ids) > 1)
    plt <- plt + facet_wrap(vars(pref_name), scales = "free_x")
  if (highlight_conc) {
    conc_data <- growth_data_bliss_nested %>%
      filter(lspci_id_1 == lspci_ids[1], lspci_id_2 == lspci_ids[2]) %>%
      chuck("data", 1)
    plt$layers <- c(
      geom_linerange(
        aes(x = drug_concentration, y = NULL), ymin = -Inf, ymax = Inf,
        color = "gray50",
        data = bind_rows(
          transmute(conc_data, lspci_id = lspci_ids[1], drug_concentration = drug_concentration_1),
          transmute(conc_data, lspci_id = lspci_ids[2], drug_concentration = drug_concentration_2),
        ) %>%
          distinct() %>%
          left_join(meta_drugs)
      ), plt$layers
    )
  }
  plt
}

plot_bliss <- function(..., fill_var = NULL) {
  args <- list(...)
  if (is.null(fill_var))
    fill_var <- "fraction_killing"
  fill_var_sym <- sym(fill_var)
  data <- args$data %>%
    mutate(across(c(drug_concentration_1, drug_concentration_2), ~fct_inseq(as.character(.x))))
  plt <- ggplot(data, aes(drug_concentration_1, drug_concentration_2, fill = !!fill_var_sym)) +
    geom_tile() +
    labs(
      title = paste(
        args$pref_name_1, "and", args$pref_name_2, "in", args$cell_name, paste0("(", args$panel_name, ")")
      ),
      x = paste(args$pref_name_1, "concentration"),
      y = paste(args$pref_name_2, "concentration"),
    ) +
    theme_minimal()
  if (fill_var %in% c("excess_bliss"))
    plt <- plt + scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-1, 1) * max(abs(data[[fill_var]])))
  else if (fill_var %in% c("gr"))
    # plt <- plt + scale_fill_viridis_c(option = "inferno", limits = c(-1.5, 1))
    plt <- plt + scale_fill_gradientn(colors = rev(adam_cmap), limits = c(NA, 1), oob = scales::squish)
  else if (str_starts(fill_var, fixed("fraction_killing")))
    plt <- plt +
      scale_fill_viridis_c(direction = -1, trans = "log2", breaks = scales::log_breaks(base = 2))
      # scale_fill_gradientn(colors = pals::parula(100), trans = "log2", breaks = scales::log_breaks(base = 2))
  else if (str_starts(fill_var, fixed("log_kills")))
    plt <- plt +
      scale_fill_viridis_c(direction = -1, limits = c(0, NA), oob = scales::squish)
  plt
}

plot_combo <- function(...) {
  args <- list(...)
  
  plot_bliss_l <- lift_dl(plot_bliss)
  
  plt <- (
      plot_bliss_l(args, fill_var = "excess_bliss") + labs(x = NULL) |
      plot_bliss_l(args, fill_var = "gr") + labs(title = NULL, x = NULL, y = NULL)
    ) /
    (
      plot_bliss_l(args, fill_var = "log_kills") + labs(title = NULL) |
      plot_bliss_l(args, fill_var = "log_kills_bliss_expected") + labs(title = NULL, y = NULL)
    ) / plot_gr(args$panel_id, args$cell_id, c(args$lspci_id_1, args$lspci_id_2), highlight_conc = TRUE) /
      plot_gr(args$panel_id, args$cell_id, c(args$lspci_id_1, args$lspci_id_2), y = log_kills, highlight_conc = TRUE)
  plt
}
x <- growth_data_bliss_nested %>% filter(lspci_id_1 == 631796, lspci_id_2 == 487093, cell_id == 16, panel_id == 9)
plt <- lift_dl(plot_combo)(x)

plt <- pmap(growth_data_bliss_nested[1, ], plot_combo)[[1]]

growth_data_bliss%>%
  filter(panel_id == 6) %>%
  arrange(desc(excess_bliss)) %>%
  head() %>%
  select(starts_with("lspci_id"), ends_with("id"))

x <- growth_data_bliss_nested %>% filter(lspci_id_1 == 4870, lspci_id_2 == 105149, cell_id == 5, panel_id == 6)
plt <- lift_dl(plot_combo)(x)
```

```{r example_combos}
plt <- pmap(filter(growth_data_bliss_nested, pref_name_1 == "GEFITINIB", pref_name_2 == "CRIZOTINIB", cell_name == "NCI-H23"), plot_combo)[[1]]
ggsave(
  file.path(dir_res, "nci_almanac_example_combo_gefi_crizo.pdf"),
  plt, width = 8, height = 9
)

plt <- pmap(filter(growth_data_bliss_nested, pref_name_1 == "DOXORUBICIN", pref_name_2 == "DOCETAXEL", cell_name == "MOLT-4"), plot_combo)[[1]]
ggsave(
  file.path(dir_res, "nci_almanac_example_combo_doxo_doce_problematic.pdf"),
  plt, width = 8, height = 9
)

plt <- pmap(filter(growth_data_bliss_nested, pref_name_1 == "CLOFARABINE", pref_name_2 == "DOCETAXEL", cell_name == "MOLT-4"), plot_combo)[[1]]
ggsave(
  file.path(dir_res, "nci_almanac_example_combo_clofa_doce.pdf"),
  plt, width = 8, height = 9
)
```

```{r bliss_vs_log_kill}

growth_data_bliss <- growth_data_bliss_raw %>%
  left_join(
    meta_cell_line
  ) %>%
  left_join(
    distinct(meta_drugs, lspci_id_1 = lspci_id, pref_name_1 = pref_name)
  ) %>%
  left_join(
    distinct(meta_drugs, lspci_id_2 = lspci_id, pref_name_2 = pref_name)
  )

calc_summary_stats <- function(sum_fn, col, name) {
  growth_data_bliss %>%
    group_by(panel_id, cell_id, lspci_id_1, lspci_id_2) %>%
    mutate(
      # Mark concentration combos that equal the requested summary stat
      is_selected = sum_fn(!!col, na.rm = TRUE) == !!col,
      excess_bliss_auc = sum(excess_bliss),
      gr_mean = mean(gr)
    ) %>%
    arrange(
      # Chose concentration combination that was selected. In case
      # of ties use highest concentration
      desc(is_selected),
      desc(drug_concentration_1 + drug_concentration_2),
      .by_group = TRUE
    ) %>%
    slice_head(n = 1) %>%
    ungroup()
}

# Define two ways we want to aggregate across concentrations:
# 1. Using the concentration combination with the lowest GR
# 2. Combo with highest log kills
growth_data_bliss_agg_nested <- tribble(
  ~sum_fn, ~col, ~name,
  min, quo(gr), "gr_min",
  max, quo(log_kills), "log_kills_max",
) %>%
  mutate(data = pmap(., calc_summary_stats))

growth_data_bliss_agg_long <- growth_data_bliss_agg_nested %>%
  select(agg_name = name, data) %>%
  unnest(data) %>%
  select(
    agg_name, panel_id, cell_id, starts_with("lspci_id"),
    starts_with("gr"), starts_with("excess_bliss"),
    starts_with("log_kills"), starts_with("fraction_killing")
  ) %>%
  pivot_longer(
    c(starts_with("gr"), starts_with("excess_bliss"),
    starts_with("log_kills"), starts_with("fraction_killing"), -gr_mean, -excess_bliss_auc)
  )

growth_data_bliss_agg <- growth_data_bliss_agg_long %>%
  mutate(name = paste(name, agg_name, sep = "_")) %>%
  select(-agg_name) %>%
  pivot_wider(
    names_from = name, values_from = value
  )

library(rasterly)
plot_scatter_rasterly <- function(
  data, x, y, out, out_dir = dir_res, x_range = NULL, y_range = NULL, ...
) {
  args <- rlang::list2(...)
  x_sym <- enquo(x)
  y_sym <- enquo(y)
  df <- data %>%
    # select(!!x_sym, !!y_sym) %>%
    drop_na()
  plt <- df %>%
    rasterly(
      mapping = aes(x = !!x_sym, y = !!y_sym),
      plot_width = args$plot_width %||% 1600, plot_height = args$plot_height %||% 1200,
      x_range = sort(x_range), y_range = sort(y_range)
    ) %>%
    rasterly_points(color = viridisLite::inferno(256)) %>%
    rasterly_build()
  sum_df <- image2data(
    plt$agg$rasterlyPoints1[[1]] %>% {
      .[rev(seq_len(nrow(.))), ]
    },
    x_range = plt$x_range, y_range = plt$y_range, background = 0
  ) %>%
    arrange(color)
  plt <- df %>%
    ggplot(aes(x, y, color = color)) +
    # stat_bin_2d(geom = "raster", bins = c(1200, 1600)) +
    geom_point(data = sum_df, shape = 16, size = 0.3, alpha = 0.8) +
    lims(x = x_range, y = y_range) +
    labs(x = args$xlab %||% rlang::as_name(x_sym), y = args$ylab %||% rlang::as_name(y_sym), color = "Count") +
    scale_color_viridis_c(trans = "log10", option = "inferno") +
    theme_minimal() +
    theme(plot.background = element_rect(fill = "white"))
  ggsave(
    file.path(out_dir, paste0(out, ".png")), plt,
    width = 10, height = 6.7
  )
  plt
}

plot_scatter_rasterly(
  growth_data_bliss, gr, excess_bliss, "nci_almanac_gr_vs_excess_bliss_scatter",
  x_range = c(1.2, -5), y_range = c(-2.5, 2.5),
  xlab = "GR", ylab = "Excess over Bliss"
)

plot_scatter_rasterly(
  growth_data_bliss, gr_bound, excess_bliss, "nci_almanac_gr_bound_vs_excess_bliss_scatter",
  x_range = c(1.2, -1.2), y_range = c(-2.5, 2.5),
  xlab = "GR bound", ylab = "Excess over Bliss"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, gr_gr_min, excess_bliss_auc, "nci_almanac_gr_min_vs_auc_scatter",
  x_range = c(1.2, -12.5), y_range = c(-10, 10),
  xlab = "GR min", ylab = "Excess over Bliss AUC"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, gr_mean, excess_bliss_auc, "nci_almanac_gr_mean_vs_auc_scatter",
  x_range = c(1.2, -6), y_range = c(-10, 10),
  xlab = "GR mean", ylab = "Excess over Bliss AUC"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, gr_gr_min, excess_bliss_gr_min, "nci_almanac_gr_min_vs_excess_bliss_gr_min",
  x_range = c(1.5, -6), y_range = c(-2.5, 2.5),
  xlab = "GR min", ylab = "Excess over Bliss at GR min"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, log_kills_log_kills_max, excess_bliss_log_kills_max, "nci_almanac_log_kills_max_vs_excess_bliss_log_kills_max",
  x_range = c(-0.2, 4.5), y_range = c(-2.5, 2.5),
  xlab = "Max log kills", ylab = "Excess over Bliss at max log kills"
)

# Plot Excess Bliss at lowest GR in combination treatment against lower GR of the two single agents
# at the selected concentration
plot_scatter_rasterly(
  growth_data_bliss_agg %>%
    # filter(
    #   gr_gr_min < 0.5
    # ) %>%
    mutate(
      lower_gr = pmax(gr_1_gr_min, gr_2_gr_min)
    ),
  lower_gr, excess_bliss_gr_min, "nci_almanac_higher_gr_single_agent_vs_excess_bliss_gr_min",
  x_range = c(1.2, -7), y_range = c(-2.5, 2.5),
  xlab = "GR of less efficacious agent at GR min", ylab = "Excess over Bliss at GR min"
)

# Plot Excess Bliss at lowest GR in combination treatment against higher GR of the two single agents
# at the selected concentration
plot_scatter_rasterly(
  growth_data_bliss_agg %>%
    # filter(
    #   gr_gr_min < 0.5
    # ) %>%
    mutate(
      lower_gr = pmin(gr_1_gr_min, gr_2_gr_min)
    ),
  lower_gr, excess_bliss_gr_min, "nci_almanac_lower_gr_single_agent_vs_excess_bliss_gr_min",
  x_range = c(1.2, -7), y_range = c(-2.5, 2.5),
  xlab = "GR of more efficacious agent at GR min", ylab = "Excess over Bliss at GR min"
)

# Plot Excess Bliss at lowest GR in combination treatment against absolute GR difference of the two single agents
# at the selected concentration
plot_scatter_rasterly(
  growth_data_bliss_agg %>%
    # filter(
    #   gr_gr_min < 0.5
    # ) %>%
    mutate(
      gr_dif_abs = abs(gr_1_gr_min - gr_2_gr_min)
    ),
  gr_dif_abs, excess_bliss_gr_min, "nci_almanac_gr_difference_single_agents_vs_excess_bliss_gr_min",
  x_range = c(0, 10), y_range = c(-2.5, 2.5),
  xlab = "Absolute GR difference of the single agents at GR min", ylab = "Excess over Bliss at GR min"
)

plot_scatter_rasterly(
  growth_data_bliss_agg %>%
    filter(
      gr_gr_min < 0.5
    ) %>%
    mutate(
      gr_dif_abs = gr_1_gr_min - gr_2_gr_min
    ),
  gr_dif_abs, excess_bliss_gr_min, "nci_almanac_gr_difference_single_agents_vs_excess_bliss_gr_min",
  x_range = c(-10, 10), y_range = c(-2.5, 2.5),
  xlab = "Absolute GR difference of the single agents at GR min", ylab = "Excess over Bliss at GR min"
)

# Plot combination log kills vs log kills of more efficacious agent
plot_scatter_rasterly(
  growth_data_bliss_agg %>%
    mutate(
      higher_log_kill = pmax(log_kills_1_gr_min, log_kills_2_gr_min)
    ),
  higher_log_kill, log_kills_gr_min, "nci_almanac_higher_log_kill_single_agent_vs_combination_log_kills_gr_min",
  x_range = c(-0.2, 3), y_range = c(-0.2, 4),
  xlab = "Log kills of more efficacious agent at GR min", ylab = "Combination log kills at GR min"
)

# Plot Bliss expectation log kills vs actual log kills
plt <- plot_scatter_rasterly(
  growth_data_bliss_agg,
  log_kills_bliss_expected_gr_min, log_kills_gr_min, "nci_almanac_bliss_expectation_vs_combination_log_kills_gr_min",
  x_range = c(-0.2, 4), y_range = c(-0.2, 4),
  xlab = "Bliss model log kills at GR min", ylab = "Combination log kills at GR min"
) +
  coord_equal() +
  geom_smooth(
    method = "lm", se=FALSE, color="black"
  )
```


```{r agg_cell}
growth_data_bliss_agg_cells <- growth_data_bliss_agg %>%
  group_by(
    lspci_id_1, lspci_id_2, 
  ) %>%
  summarize(across(where(is.numeric), mean), .groups = "drop")

plot_scatter_rasterly(
  growth_data_bliss_agg_cells, gr_gr_min, excess_bliss_auc, "nci_almanac_cells_agg_gr_min_vs_auc_scatter",
  x_range = c(1.2, -5), y_range = c(-6, 6),
  xlab = "GR min", ylab = "Excess over Bliss AUC"
)
```

```{r agg_cell}
growth_data_bliss_agg_cells <- growth_data_bliss_agg %>%
  group_by(
    lspci_id_1, lspci_id_2, 
  ) %>%
  summarize(across(where(is.numeric), mean), .groups = "drop")

plot_scatter_rasterly(
  growth_data_bliss_agg_cells, gr_gr_min, excess_bliss_auc, "nci_almanac_cells_agg_gr_min_vs_auc_scatter",
  x_range = c(1.2, -5), y_range = c(-6, 6),
  xlab = "GR min", ylab = "Excess over Bliss AUC"
)

plot_scatter_rasterly(
  growth_data_bliss_agg_cells, gr_gr_min, excess_bliss_gr_min, "nci_almanac_cells_agg_gr_min_vs_excess_bliss_gr_min_scatter",
  x_range = c(1.2, -5), y_range = c(-1.5, 1.5),
  xlab = "GR min", ylab = "Excess over Bliss at GR min"
)

plot_scatter_rasterly(
  growth_data_bliss_agg_cells, gr_mean, excess_bliss_auc, "nci_almanac_cells_agg_gr_mean_vs_auc_scatter",
  x_range = c(1.2, -1), y_range = c(-6, 6),
  xlab = "GR mean", ylab = "Excess over Bliss AUC"
)
```


```{r variance}
p <- growth_data_bliss_agg %>%
  left_join(meta_cell_line) %>%
  # mutate(across(c(panel_id, cell_id), as.factor)) %>%
  ggplot(
    aes(panel_name, gr_bound_gr_min)
  ) +
    geom_violin() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

p <- growth_data_bliss_agg %>%
  left_join(meta_cell_line) %>%
  ggplot(
    aes(1, gr_bound_gr_min)
  ) +
    geom_violin()

growth_data_bliss_agg_var <- tribble(
  ~var_across, ~group_vars,
  "combination", list(sym("lspci_id_1"), sym("lspci_id_2")),
  "cell_line", list(sym("panel_id"), sym("cell_id"))
)
growth_data_bliss_agg_var_data <- map(
  growth_data_bliss_agg_var$group_vars,
  function(g) {
    growth_data_bliss_agg %>%
      group_by(!!!g) %>%
      summarize(across(where(is.numeric), sd), .groups = "drop")
  }
)
growth_data_bliss_agg_var <- growth_data_bliss_agg_var %>%
  mutate(data = growth_data_bliss_agg_var_data) %>%
  select(-group_vars) %>%
  unnest(data)

p <- growth_data_bliss_agg_var %>%
  ggplot(aes(var_across, gr_bound_gr_min)) +
    # geom_violin() +
    geom_quasirandom() +
    labs(x = NULL, y = "Standard deviation GR bound")
ggsave(
  file.path(dir_res, "standard_deviation_beeswarm_gr_bound_gr_min.pdf"),
  p, width = 3, height = 5
)

p <- growth_data_bliss_agg_var %>%
  ggplot(aes(var_across, fraction_killing_gr_min)) +
    # geom_violin() +
    geom_quasirandom() +
    labs(x = NULL, y = "Standard deviation fraction killing")
ggsave(
  file.path(dir_res, "standard_deviation_beeswarm_fraction_killing_gr_min.pdf"),
  p, width = 3, height = 5
)
```


```{r}
## By cell line
dir_res_by_cell <- file.path(dir_res, "by_cell_line")
dir.create(dir_res_by_cell, showWarnings = FALSE)

x <- growth_data_bliss_auc %>%
  filter(
    panel_id == 9, cell_id == 18,
    log_kills_max > 1.5, excess_bliss_log_kills_max > 0
  )

y <- growth_data_bliss %>%
  filter(
    panel_id == 9, cell_id == 18,
    log_kills > 1.5, excess_bliss > 0,
    lspci_id_1 == 55206
  )

scatter_plots_per_cell_line <- growth_data_bliss %>%
  distinct(panel_id, cell_id) %>%
  pmap(
    function(panel_id, cell_id) {
      cell_info <- meta_cell_line %>%
        mutate(across(c(cell_name, panel_name), str_replace_all, r"([^\w])", "_")) %>%
        filter(cell_id == !!cell_id, panel_id == !!panel_id) %>% {
          paste(.$panel_name, .$cell_name, sep = "_")
        }
      
      print(cell_info)
      
      plots <- list()
      
      plots[["gr_vs_excess_bliss"]] <- plot_scatter_rasterly(
        filter(growth_data_bliss, cell_id == !!cell_id, panel_id == !!panel_id),
        gr, excess_bliss, paste0("nci_almanac_", cell_info, "_gr_vs_excess_bliss_scatter"),
        x_range = c(1.5, -3), y_range = c(-1, 1),
        xlab = "GR", ylab = "Excess over Bliss",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      df_auc <- filter(growth_data_bliss_auc, cell_id == !!cell_id, panel_id == !!panel_id)
      plots[["gr_min_vs_auc"]] <- plot_scatter_rasterly(
        df_auc, gr_min, AUC, paste0("nci_almanac_", cell_info, "_gr_min_vs_auc_scatter"),
        x_range = c(1.5, -3), y_range = c(-5, 5),
        xlab = "GR min", ylab = "Excess over Bliss AUC",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["gr_mid_vs_excess_bliss_gr_mid"]] <- plot_scatter_rasterly(
        df_auc, gr_mid, excess_bliss_gr_mid, paste0("nci_almanac_", cell_info, "_gr_mid_vs_excess_bliss_gr_mid_scatter"),
        x_range = c(1.5, -3), y_range = c(-1, 1),
        xlab = "GR mid concentration", ylab = "Excess over Bliss mid concentration",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["gr_mean_vs_auc"]] <- plot_scatter_rasterly(
        df_auc, gr_mean, AUC, paste0("nci_almanac_", cell_info, "_gr_mean_vs_auc_scatter"),
        x_range = c(1.5, -3), y_range = c(-5, 5),
        xlab = "GR mean", ylab = "Excess over Bliss AUC",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["gr_min_vs_auc"]] <- plot_scatter_rasterly(
        df_auc, gr_min, AUC, paste0("nci_almanac_", cell_info, "_gr_min_vs_auc_scatter"),
        x_range = c(1.5, -3), y_range = c(-5, 5),
        xlab = "GR min", ylab = "Excess over Bliss AUC",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
       plots[["gr_min_vs_excess_bliss_gr_min"]] <- plot_scatter_rasterly(
        df_auc, gr_min, excess_bliss_gr_min, paste0("nci_almanac_", cell_info, "_gr_min_vs_excess_bliss_gr_min"),
        x_range = c(1.5, -3), y_range = c(-1, 1),
        xlab = "GR min", ylab = "Excess over Bliss at GR min",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["log_kills_max_vs_excess_bliss_log_kills_max"]] <- plot_scatter_rasterly(
        df_auc, log_kills_max, excess_bliss_log_kills_max, paste0("nci_almanac_", cell_info, "_log_kills_max_vs_excess_bliss_log_kills_max"),
        x_range = c(-0.2, 3), y_range = c(-1, 1),
        xlab = "Max log kills", ylab = "Excess over Bliss at max log kills",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots
    }
  )


```

```{r response_rate}


rr_bliss_cell_agg <- tibble(response_threshold = seq(0.2, 1, by = 0.05)) %>%
  mutate(
    data = map(
      response_threshold,
      function(response_threshold) {
        growth_data_bliss_agg_long %>%
          select(-gr_mean, -excess_bliss_auc) %>%
          filter(str_starts(name, fixed("fraction_killing"))) %>%
          pivot_wider(names_from = name, values_from = value) %>%
          group_by(
            across(c(agg_name, starts_with("lspci_id")))
          ) %>%
          summarize(
            across(
              c(starts_with("fraction_killing")),
              ~sum(.x > response_threshold) / length(.x)
            ), .groups = "drop"
          )
      }
    )
  ) %>%
  unnest(data) %>%
  compute_bliss()

plot_scatter_rasterly(
  filter(rr_bliss_cell_agg, response_threshold == 0.2, agg_name == "gr_min"),
  fraction_killing, excess_bliss,
  "nci_almanac_cell_agg_response_rate_0.5_gr_min_vs_excess_independence_gr_min",
  x_range = c(0, 1), y_range = c(-1, 1),
  xlab = "Response rate at GR min", ylab = "Excess over independence at GR min"
)

plt <- growth_data_bliss_cell_agg %>%
  ggplot(
    aes(fraction_killing_gr_min, excess_bliss_gr_min)
  ) +
    geom_point() +
    facet_wrap(~response_threshold) +
    labs(x = "Response rate at GR min", y = "Excess over independence at GR min") +
    # lims(x = c(0, 1), y = c(-1, 1)) +
    theme_minimal()

plot_scatter_rasterly(
  filter(growth_data_bliss_cell_agg, panel_id == 1, response_threshold == 0.95), fraction_killing_gr_min, excess_bliss_gr_min, "nci_almanac_cell_agg_response_rate_gr_min_vs_excess_independence_gr_min",
  x_range = c(0, 1), y_range = c(-1, 1),
  xlab = "Response rate at GR min", ylab = "Excess over independence at GR min"
)

```


```{r proportion_synergistic}
synergy_threshold <- 1

proportion_synergistic <- growth_data_bliss_agg %>%
  mutate(
    gr_bin = cut(gr_gr_min, c(-Inf, seq(min(gr_gr_min), max(gr_gr_min), length.out = 20), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(gr_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss_gr_min > synergy_threshold) / n(),
    proportion_synergistic_combos = sum(excess_bliss_gr_min > synergy_threshold) / sum(.$excess_bliss_gr_min > synergy_threshold),
    .groups = "drop"
  )

plt <- proportion_synergistic %>%
  pivot_longer(starts_with("proportion"), names_to = "metric", values_to = "proportion") %>%
  ggplot(aes(gr_bin, proportion)) +
    geom_col() +
    facet_wrap(~metric, scales = "free_y", ncol = 1) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
ggsave(
  file.path(dir_res, "nci_almanac_proportion_synergistic_at_gr_min.pdf"),
  plt, width = 4, height = 5
)

x <- growth_data_bliss_agg %>%
  filter(gr_gr_min < -7, gr_gr_min > -8, excess_bliss_gr_min > 1) %>%
  arrange(panel_id, cell_id, lspci_id_1, lspci_id_2)
xx <- distinct(x, panel_id, cell_id, lspci_id_1, lspci_id_2)

y <- growth_data_bliss %>%
  semi_join(x)

z <- growth_data_bliss_nested %>%
  semi_join(x) %>%
  arrange(panel_id, cell_id, lspci_id_1, lspci_id_2)
zz <- distinct(z, panel_id, cell_id, lspci_id_1, lspci_id_2)

  filter(lspci_id_1 == 631796, lspci_id_2 == 487093, cell_id == 16, panel_id == 9)
plt <- lift_dl(plot_combo)(x)
z_plots <- z %>%
  pmap(plot_combo)

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(gr_bin) %>%
  summarize(
    proportion_synergistic_combos_per_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  ggplot(aes(gr_bin, proportion_synergistic_combos_per_bin)) +
    geom_col()

plt <- growth_data_bliss %>%
  mutate(
    log_kills_bin = cut(log_kills, c(-Inf, seq(0.5, 9.5, by = 0.5), Inf), ordered_result = TRUE)
  ) %>%
  group_by(log_kills_bin) %>%
  summarize(
    proportion_synergistic_combos_per_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  ggplot(aes(log_kills_bin, proportion_synergistic_combos_per_bin)) +
    geom_col()

plt <- growth_data_bliss_auc %>%
  mutate(
    gr_mean_bin = cut(gr_mean, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_mean_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss_mean > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss_mean > log10(2)) / sum(.$excess_bliss_mean > log10(2)),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_mean_bin, proportion_synergistic_combos_in_bin)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss_auc %>%
  mutate(
    gr_mean_bin = cut(gr_mean, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_mean_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss_mean > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss_mean > log10(2)) / sum(.$excess_bliss_mean > log10(2)),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_mean_bin, proportion_synergistic_combos_in_bin)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_bin) %>%
  filter(n() > 10) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_bin, proportion_synergistic_combos_in_bin)) +
    # geom_crossbar(stat = "summary", color = "gray40") +
    # ggbeeswarm::geom_quasirandom(width = 0.2) +
    geom_line(aes(group = cell_id)) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_bin) %>%
  filter(n() > 10) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    excess_bliss_mean = mean(excess_bliss),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_bin, excess_bliss_mean)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    # ggbeeswarm::geom_quasirandom(width = 0.2) +
    geom_line(aes(group = cell_id)) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_bin) %>%
  filter(n() > 10) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    excess_bliss_mean = mean(excess_bliss),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_bin, excess_bliss_mean)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    # ggbeeswarm::geom_quasirandom(width = 0.2) +
    geom_line(aes(group = cell_id)) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

```



```{r umap}
library(umap)

get_mat <- function(df, var) {
  var_sym <- enquo(var)
  df_long <- df %>%
    transmute(
      panel_cell_id = paste(panel_id, cell_id, sep = "_"),
      lspci_id_combo = paste(lspci_id_1, lspci_id_2, sep = "_"),
      !!var_sym
    )
  df_wide <- df_long %>%
    pivot_wider(names_from = lspci_id_combo, values_from = !!var_sym)
  mat <- df_wide %>%
    column_to_rownames("panel_cell_id") %>%
    as.matrix()
  mat
}

impute_mat <- function(mat, ...) {
  # imp_formula <- reformulate(paste0("`", colnames(mat), "`"))
  # Hmisc::aregImpute(
  #   imp_formula, data = as.data.frame(mat), n.impute = 5, ...
  # )
  VIM::kNN(mat, ...)
}

gr_min_mat <- get_mat(growth_data_bliss_auc, gr_min)
gr_min_mat_imp <- impute_mat(gr_min_mat)
  

umap_gr_min <- umap(
  growth_data_bliss_auc %>%
    transmute(panel_cell_id = paste(panel_id, cell_id, sep = ":"), lspci_id_1, lspci_id_2, gr_min) %>%
    pivot_wider(names_from = panel_cell_id, values_from = gr_min) %>%
    select(-starts_with("lspci_id")) %>%
    as.matrix()
)
```

* Net killing vs Excess over Bliss, using all concentrations
* Or using AUC, adding up all values
* Stratify by cell line
* Use GR value for x-axis
* Range of Excess over Bliss across cell line
* Plot minimum GR vs excess killing at the same doses

* UMAP of AUC values per cell line
