---
title: "Plot NCI data"
author: "Clemens Hug"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(here)
library(ggbeeswarm)
library(patchwork)

dir_data <- here("data")
dir.create(dir_data, showWarnings = FALSE)
dir_res <- here("nci")
dir.create(dir_res, showWarnings = FALSE)

synapser::synLogin()
syn <- synExtra::synDownloader(dir_data)
```

## Load data from Synapse

```{r load}
growth_data_raw <- syn("syn25882596") %>%
  read_csv()
meta_cell_line <- syn("syn25882595") %>%
  read_csv()
meta_drugs <- syn("syn25882594") %>%
  read_csv()
```

## Format data for plotting

PercentGrowth    NUMBER           percent growth of drug combination
against cell line, using time zero in
calculation
PercentGrowthNoTZ NUMBER          percent growth without time zero
TestValue         NUMBER          test mean optical density
ControlValue      NUMBER          vehicle control mean optical density
TZValue           NUMBER          time zero mean optical density
ExpectedGrowth    NUMBER          expected percent growth for combination,
based on values for individual drugs
Score             NUMBER(4)       score, for combination records

```{r gr}
compute_metrics <- function(data) {
  mutate(
    data,
    fraction_killing = 1 - (TESTVALUE / CONTROLVALUE),
    log_killing = log10(CONTROLVALUE / TESTVALUE),
    gr = log2(TESTVALUE / TZVALUE) / log2(CONTROLVALUE / TZVALUE),
    gr_bound = 2**gr - 1
  )
}

growth_data <- growth_data_raw %>%
  compute_metrics()

growth_data_single_dose_all <- growth_data %>%
  filter(is.na(lspci_id_2)) %>%
  select(
    panel_id, cell_id,
    lspci_id = lspci_id_1,
    drug_concentration = drug_concentration_1,
    ends_with("killing"),
    ends_with("VALUE"),
    starts_with("gr")
  )

growth_data_single_dose_agg <- growth_data_single_dose_all %>%
  group_by(panel_id, cell_id, lspci_id, drug_concentration) %>%
  summarize(
    across(ends_with("VALUE"), mean),
    .groups = "drop"
  ) %>%
  compute_metrics()
  
growth_data_bliss <- growth_data %>%
  filter(!is.na(lspci_id_2)) %>%
  left_join(
    growth_data_single_dose_agg %>%
      select(
        panel_id, cell_id,
        lspci_id_1 = lspci_id, drug_concentration_1 = drug_concentration,
        fraction_killing_1 = fraction_killing
      )
  ) %>%
  left_join(
    growth_data_single_dose_agg %>%
      select(
        panel_id, cell_id,
        lspci_id_2 = lspci_id, drug_concentration_2 = drug_concentration,
        fraction_killing_2 = fraction_killing
      )
  ) %>%
  mutate(
    fraction_killing_bliss_expected = fraction_killing_1 + fraction_killing_2 - fraction_killing_1 * fraction_killing_2,
    excess_killing = log10(1 - fraction_killing_bliss_expected) - log10(1 - fraction_killing)
  )
```

```{r gr_descriptive}
plt <- growth_data_gr %>%
  ggplot(aes(gr)) +
     geom_linerange(
      aes(q, ymin = -0.3, ymax = 0),
      data = ~quantile(.x$gr, seq(.1, 0.9, by = 0.1)) %>%
        enframe("quantile", "q")
    ) +
    geom_histogram(aes(y = stat(density))) +
    geom_density(color = "red") +
    lims(x = c(-2, NA))

ggsave(
  file.path(dir_res, "desc_gr_histogram.pdf"), plt,
  width = 4, height = 2
)

```


```{r plot_bliss}
growth_data_bliss_nested <- growth_data_bliss %>%
  group_nest(
    panel_id, cell_id, lspci_id_1, lspci_id_2
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_1 = lspci_id, pref_name_1 = pref_name)
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_2 = lspci_id, pref_name_2 = pref_name)
  )

adam_cmap <- c(
  "#FFDA64", "#FFC952", "#FAB534", "#FCAB24", "#F79820", "#F48221", "#F4751C",
  "#EB6820", "#D95330", "#D04445", "#B93B51", "#953565", "#78287E", "#522B6E", "#22152F"
)

plot_gr <- function(panel_id, cell_id, lspci_ids, highlight_conc = FALSE) {
  data_agg <- growth_data_single_dose_agg %>%
    filter(panel_id == !!panel_id, cell_id == !!cell_id, lspci_id %in% lspci_ids) %>%
    left_join(meta_drugs)
  data_all <- growth_data_single_dose_all %>%
    filter(panel_id == !!panel_id, cell_id == !!cell_id, lspci_id %in% lspci_ids) %>%
    left_join(meta_drugs)
  plt <- ggplot(data_agg, aes(drug_concentration, gr)) +
    geom_line(color = "gray") +
    geom_beeswarm(data = data_all, groupOnX = TRUE) +
    scale_x_log10() +
    labs(x = "Concentration") +
    theme_minimal()
  if (length(lspci_ids) > 1)
    plt <- plt + facet_wrap(vars(pref_name), scales = "free_x")
  if (highlight_conc) {
    conc_data <- growth_data_bliss_nested %>%
      filter(lspci_id_1 == lspci_ids[1], lspci_id_2 == lspci_ids[2]) %>%
      chuck("data", 1)
    plt <- plt +
      geom_linerange(
        aes(x = drug_concentration, y = NULL), ymin = -Inf, ymax = Inf,
        data = bind_rows(
          transmute(conc_data, lspci_id = lspci_ids[1], drug_concentration = drug_concentration_1),
          transmute(conc_data, lspci_id = lspci_ids[2], drug_concentration = drug_concentration_2),
        ) %>%
          distinct() %>%
          left_join(meta_drugs)
      )
  }
  plt
}

plot_bliss <- function(..., fill_var = NULL) {
  args <- list(...)
  if (is.null(fill_var))
    fill_var <- "fraction_killing"
  fill_var_sym <- sym(fill_var)
  data <- args$data[[1]] %>%
    mutate(across(c(drug_concentration_1, drug_concentration_2), ~fct_inseq(as.character(.x))))
  plt <- ggplot(data, aes(drug_concentration_1, drug_concentration_2, fill = !!fill_var_sym)) +
    geom_tile() +
    labs(
      title = paste(
        args$pref_name_1, "and", args$pref_name_2, "in", args$cell_name, paste0("(", args$panel_name, ")")
      ),
      x = paste(args$pref_name_1, "concentration"),
      y = paste(args$pref_name_2, "concentration"),
    ) +
    theme_minimal()
  if (fill_var %in% c("excess_killing"))
    plt <- plt + scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-1, 1) * max(abs(data[[fill_var]])))
  else if (fill_var %in% c("gr"))
    # plt <- plt + scale_fill_viridis_c(option = "inferno", limits = c(-1.5, 1))
    plt <- plt + scale_fill_gradientn(colors = rev(adam_cmap), limits = c(-1.5, 1))
  else if (str_starts(fill_var, fixed("fraction_killing")) | fill_var %in% c("log_killing"))
    plt <- plt +
      scale_fill_viridis_c(direction = -1, trans = "log2", breaks = scales::log_breaks(base = 2))
      # scale_fill_gradientn(colors = pals::parula(100), trans = "log2", breaks = scales::log_breaks(base = 2))
  plt
}

plot_combo <- function(...) {
  args <- list(...)
  
  plot_bliss_l <- lift_dl(plot_bliss)
  
  plt <- (
      plot_bliss_l(args, fill_var = "excess_killing") + labs(x = NULL) |
      plot_bliss_l(args, fill_var = "gr") + labs(title = NULL, x = NULL, y = NULL)
    ) /
    (
      plot_bliss_l(args, fill_var = "fraction_killing") + labs(title = NULL) |
      plot_bliss_l(args, fill_var = "fraction_killing_bliss_expected") + labs(title = NULL, y = NULL)
    ) / plot_gr(args$panel_id, args$cell_id, c(args$lspci_id_1, args$lspci_id_2), highlight_conc = TRUE)
  plt
}
x <- growth_data_bliss_nested %>% filter(lspci_id_1 == 487093, lspci_id_2 == 631796, cell_id == 16, panel_id == 9)
plt <- lift_dl(plot_combo)(x)

plt <- lift_dl(plot_combo)(growth_data_bliss_nested[545, ])
plt

```

```{r bliss_vs_log_kill}

growth_data_bliss <- growth_data_bliss %>%
  left_join(
    meta_cell_line
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_1 = lspci_id, pref_name_1 = pref_name)
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_2 = lspci_id, pref_name_2 = pref_name)
  )

library(rasterly)

plt <- growth_data_bliss %>% 
  drop_na(gr, excess_killing) %>%
  rasterly(mapping = aes(x = gr, y = excess_killing), x_range = c(-3, 1.5), y_range = c(-1.5, 1.5)) %>% 
  rasterly_points(color = viridisLite::inferno(256))

x <- plot(plt, xlab = "GR", ylab = "Excess over Bliss")

library(reticulate)

callr::r(function(df){
  library(reticulate)
  Sys.setenv(PYTHONPATH = "", PYTHONHOME = "")
  use_condaenv("datashader", conda = "/Users/chug/opt/miniconda3/bin/conda", required = TRUE)
  ds <- import("datashader")
  cc <- import("colorcet")
  cvs <- ds$Canvas(plot_width = 1600L, plot_height = 1200L, x_range = c(-3, 1.5), y_range = c(-1.5, 1.5))
  agg <- cvs$points(df, "gr", "excess_killing")
  img <- ds$tf$set_background(ds$tf$shade(agg, how = "eq_hist", cmap = cc$fire), "black")
  ds$utils$export_image(img = img, filename = 'gr_excess_killing.png', fmt = ".png", background = NULL)
}, show = TRUE, args = list(df = select(growth_data_bliss, gr, excess_killing) %>% drop_na()))

```



* Net killing vs delta bliss, using all concentrations
* Or using AUC, adding up all values
* Stratify by cell line
* Use GR value for x-axis
* Range of delta bliss across cell line

* UMAP of AUC values per cell line
