---
title: "Plot NCI data"
author: "Clemens Hug"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(here)
library(ggbeeswarm)
library(patchwork)

dir_data <- here("data")
dir.create(dir_data, showWarnings = FALSE)
dir_res <- here("nci")
dir.create(dir_res, showWarnings = FALSE)

synapser::synLogin()
syn <- synExtra::synDownloader(dir_data, ifcollision = "overwrite.local")
```

## Load data from Synapse

```{r load}
growth_data_raw <- syn("syn25882596") %>%
  read_csv()
meta_cell_line <- syn("syn25882595") %>%
  read_csv()
meta_drugs <- syn("syn25882594") %>%
  read_csv()
```

## Format data for plotting

PercentGrowth    NUMBER           percent growth of drug combination
against cell line, using time zero in
calculation
PercentGrowthNoTZ NUMBER          percent growth without time zero
TestValue         NUMBER          test mean optical density
ControlValue      NUMBER          vehicle control mean optical density
TZValue           NUMBER          time zero mean optical density
ExpectedGrowth    NUMBER          expected percent growth for combination,
based on values for individual drugs
Score             NUMBER(4)       score, for combination records

```{r qc}
StatPercentileX <- ggproto("StatPercentileX", Stat,
  compute_group = function(data, scales, probs) {
    percentiles <- quantile(data$x, probs=probs)
    data.frame(xintercept=percentiles, x=percentiles)
    },
  required_aes = c("x")
)

stat_percentile_x <- function(mapping = NULL, data = NULL, geom = "vline",
                              position = "identity", na.rm = FALSE,
                              show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    stat = StatPercentileX, data = data, mapping = mapping, geom = geom, 
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

diagnostic_histogram <- function(df, x) {
  x_sym <- enquo(x)
  df %>%
    ggplot(aes(!!x_sym)) +
      stat_percentile_x(geom = "linerange", ymin = -0.05, ymax = 0, probs = seq(.1, 0.9, by = 0.1), color = "red") +
      # geom_linerange(
      #   aes(q, ymin = -0.05, ymax = 0),
      #   data = ~quantile(.x[[rlang::as_name(x_sym)]], seq(.1, 0.9, by = 0.1), na.rm = TRUE) %>%
      #     enframe("quantile", "q"),
      #   color = "red"
      # ) +
      geom_histogram(aes(y = stat(ndensity)), bins = 50)
      # geom_density(color = "red")
}

plt <- growth_data_raw %>%
  diagnostic_histogram(CONTROLVALUE) +
  scale_x_log10() +
  facet_wrap(vars(study_location))
ggsave(file.path(dir_res, "nci_almanac_CONTROLVALUE_histogram.pdf"), plt, width = 9, height = 3)

plt <- growth_data_raw %>%
  diagnostic_histogram(TESTVALUE) +
  scale_x_log10() +
  facet_wrap(vars(study_location))
ggsave(file.path(dir_res, "nci_almanac_TESTVALUE_histogram.pdf"), plt, width = 9, height = 3)

plt <- growth_data_raw %>%
  diagnostic_histogram(TZVALUE) +
  scale_x_log10() +
  facet_wrap(vars(study_location))
ggsave(file.path(dir_res, "nci_almanac_TZVALUE_histogram.pdf"), plt, width = 9, height = 3)

plt <- growth_data_raw %>%
  count(study_location, measurement_type = if_else(is.na(lspci_id_2), "single_agent", "combination")) %>%
  ggplot(aes(study_location, n, fill = measurement_type)) +
    geom_col(position = "dodge")
ggsave(file.path(dir_res, "nci_almanac_study_location_counts.pdf"), plt, width = 5, height = 3)

growth_data_raw %>%
  drop_na(lspci_id_2) %>%
  group_by(
    panel_id, cell_id, lspci_id_1, lspci_id_2, drug_concentration_1, drug_concentration_2
  ) %>%
  arrange(.by_group = TRUE) %>%
  filter(n() > 1)

growth_data_raw %>%
  group_by(
    lspci_id_1, measurement_type = if_else(is.na(lspci_id_2), "single_agent", "combination")
  ) %>%
  summarize(n_locations = length(unique(study_location)), measurement_types = .groups = "drop")
```



```{r gr}
compute_metrics <- function(data) {
  mutate(
    data,
    # fraction_killing = pmax(0, 1 - (TESTVALUE / CONTROLVALUE)),
    # log_kills = pmax(0, log10(CONTROLVALUE / TESTVALUE)),
    fraction_killing = 1 - (TESTVALUE / CONTROLVALUE),
    log_kills = log10(CONTROLVALUE / TESTVALUE),
    gr = log2(TESTVALUE / TZVALUE) / log2(CONTROLVALUE / TZVALUE),
    gr_bound = 2**gr - 1
  )
}

growth_data <- growth_data_raw %>%
  # # Remove data with extremely low control or time zero counts
  # # Actually, even the low values seem to be fine. They result from
  # # a different protocol being used by one of the study centers
  # # So we only have to filter for negative values
  # filter(
  #   CONTROLVALUE > 1e3,
  #   TZVALUE > 1e3
  # ) %>%
  filter(TESTVALUE > 0) %>%
  compute_metrics()

growth_data_single_dose_all <- growth_data %>%
  filter(is.na(lspci_id_2)) %>%
  select(
    panel_id, cell_id,
    lspci_id = lspci_id_1,
    drug_concentration = drug_concentration_1,
    ends_with("killing"),
    ends_with("VALUE"),
    starts_with("gr"),
    contains("kills")
  )

growth_data_single_dose_agg <- growth_data_single_dose_all %>%
  group_by(panel_id, cell_id, lspci_id, drug_concentration) %>%
  summarize(
    across(ends_with("VALUE"), mean),
    .groups = "drop"
  ) %>%
  compute_metrics()
  
growth_data_bliss_raw <- growth_data %>%
  filter(!is.na(lspci_id_2)) %>%
  left_join(
    growth_data_single_dose_agg %>%
      select(
        panel_id, cell_id,
        lspci_id_1 = lspci_id, drug_concentration_1 = drug_concentration,
        fraction_killing_1 = fraction_killing
      )
  ) %>%
  left_join(
    growth_data_single_dose_agg %>%
      select(
        panel_id, cell_id,
        lspci_id_2 = lspci_id, drug_concentration_2 = drug_concentration,
        fraction_killing_2 = fraction_killing
      )
  ) %>%
  mutate(
    fraction_killing_bliss_expected = fraction_killing_1 + fraction_killing_2 - fraction_killing_1 * fraction_killing_2,
    excess_bliss = log10(1 - fraction_killing_bliss_expected) - log10(1 - fraction_killing),
    log_kills_bliss_expected = log10(1 / (1 - fraction_killing_bliss_expected))
  )

plt <- growth_data %>%
  diagnostic_histogram(gr) +
    lims(x = c(-2, NA)) +
    facet_wrap(vars(study_location))

ggsave(
  file.path(dir_res, "nci_almanac_gr_histogram.pdf"), plt,
  width = 6, height = 3
)
```

```{r plot_bliss}
growth_data_bliss_nested <- growth_data_bliss_raw %>%
  group_nest(
    panel_id, cell_id, lspci_id_1, lspci_id_2
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_1 = lspci_id, pref_name_1 = pref_name)
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_2 = lspci_id, pref_name_2 = pref_name)
  )

adam_cmap <- c(
  "#FFDA64", "#FFC952", "#FAB534", "#FCAB24", "#F79820", "#F48221", "#F4751C",
  "#EB6820", "#D95330", "#D04445", "#B93B51", "#953565", "#78287E", "#522B6E", "#22152F"
)

plot_gr <- function(panel_id, cell_id, lspci_ids, y = gr, highlight_conc = FALSE) {
  y_sym <- enquo(y)
  data_agg <- growth_data_single_dose_agg %>%
    filter(panel_id == !!panel_id, cell_id == !!cell_id, lspci_id %in% lspci_ids) %>%
    left_join(meta_drugs)
  data_all <- growth_data_single_dose_all %>%
    filter(panel_id == !!panel_id, cell_id == !!cell_id, lspci_id %in% lspci_ids) %>%
    left_join(meta_drugs)
  plt <- ggplot(data_agg, aes(drug_concentration, !!y_sym)) +
    geom_line(color = "gray") +
    geom_beeswarm(data = data_all, groupOnX = TRUE) +
    scale_x_log10() +
    labs(x = "Concentration") +
    theme_minimal()
  if (length(lspci_ids) > 1)
    plt <- plt + facet_wrap(vars(pref_name), scales = "free_x")
  if (highlight_conc) {
    conc_data <- growth_data_bliss_nested %>%
      filter(lspci_id_1 == lspci_ids[1], lspci_id_2 == lspci_ids[2]) %>%
      chuck("data", 1)
    plt$layers <- c(
      geom_linerange(
        aes(x = drug_concentration, y = NULL), ymin = -Inf, ymax = Inf,
        color = "gray50",
        data = bind_rows(
          transmute(conc_data, lspci_id = lspci_ids[1], drug_concentration = drug_concentration_1),
          transmute(conc_data, lspci_id = lspci_ids[2], drug_concentration = drug_concentration_2),
        ) %>%
          distinct() %>%
          left_join(meta_drugs)
      ), plt$layers
    )
  }
  plt
}

plot_bliss <- function(..., fill_var = NULL) {
  args <- list(...)
  if (is.null(fill_var))
    fill_var <- "fraction_killing"
  fill_var_sym <- sym(fill_var)
  data <- args$data[[1]] %>%
    mutate(across(c(drug_concentration_1, drug_concentration_2), ~fct_inseq(as.character(.x))))
  plt <- ggplot(data, aes(drug_concentration_1, drug_concentration_2, fill = !!fill_var_sym)) +
    geom_tile() +
    labs(
      title = paste(
        args$pref_name_1, "and", args$pref_name_2, "in", args$cell_name, paste0("(", args$panel_name, ")")
      ),
      x = paste(args$pref_name_1, "concentration"),
      y = paste(args$pref_name_2, "concentration"),
    ) +
    theme_minimal()
  if (fill_var %in% c("excess_bliss"))
    plt <- plt + scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-1, 1) * max(abs(data[[fill_var]])))
  else if (fill_var %in% c("gr"))
    # plt <- plt + scale_fill_viridis_c(option = "inferno", limits = c(-1.5, 1))
    plt <- plt + scale_fill_gradientn(colors = rev(adam_cmap), limits = c(NA, 1), oob = scales::squish)
  else if (str_starts(fill_var, fixed("fraction_killing")))
    plt <- plt +
      scale_fill_viridis_c(direction = -1, trans = "log2", breaks = scales::log_breaks(base = 2))
      # scale_fill_gradientn(colors = pals::parula(100), trans = "log2", breaks = scales::log_breaks(base = 2))
  else if (str_starts(fill_var, fixed("log_kills")))
    plt <- plt +
      scale_fill_viridis_c(direction = -1, limits = c(0, NA), oob = scales::squish)
  plt
}

plot_combo <- function(...) {
  args <- list(...)
  
  plot_bliss_l <- lift_dl(plot_bliss)
  
  plt <- (
      plot_bliss_l(args, fill_var = "excess_bliss") + labs(x = NULL) |
      plot_bliss_l(args, fill_var = "gr") + labs(title = NULL, x = NULL, y = NULL)
    ) /
    (
      plot_bliss_l(args, fill_var = "log_kills") + labs(title = NULL) |
      plot_bliss_l(args, fill_var = "log_kills_bliss_expected") + labs(title = NULL, y = NULL)
    ) / plot_gr(args$panel_id, args$cell_id, c(args$lspci_id_1, args$lspci_id_2), highlight_conc = TRUE) /
      plot_gr(args$panel_id, args$cell_id, c(args$lspci_id_1, args$lspci_id_2), y = log_kills, highlight_conc = TRUE)
  plt
}
x <- growth_data_bliss_nested %>% filter(lspci_id_1 == 487093, lspci_id_2 == 631796, cell_id == 16, panel_id == 9)
plt <- lift_dl(plot_combo)(x)

plt <- lift_dl(plot_combo)(growth_data_bliss_nested[999, ])
plt

growth_data_bliss%>%
  filter(panel_id == 6) %>%
  arrange(desc(excess_bliss)) %>%
  head() %>%
  select(starts_with("lspci_id"), ends_with("id"))

x <- growth_data_bliss_nested %>% filter(lspci_id_1 == 4870, lspci_id_2 == 105149, cell_id == 5, panel_id == 6)
plt <- lift_dl(plot_combo)(x)
```

```{r bliss_vs_log_kill}


growth_data_bliss <- growth_data_bliss_raw %>%
  left_join(
    meta_cell_line
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_1 = lspci_id, pref_name_1 = pref_name)
  ) %>%
  left_join(
    select(meta_drugs, lspci_id_2 = lspci_id, pref_name_2 = pref_name)
  )

 calc_summary_stats <- function(sum_fn, col, name) {
  growth_data_bliss %>%
    group_by(panel_id, cell_id, lspci_id_1, lspci_id_2) %>%
    mutate(
      # Mark concentration combos that equal the requested summary stat
      is_selected = sum_fn(!!col, na.rm = TRUE) == !!col,
      excess_bliss_auc = sum(excess_bliss),
      gr_mean = mean(gr)
    ) %>%
    arrange(
      # Chose concentration combination that was selected. In case
      # of ties use highest concentration
      desc(is_selected),
      desc(drug_concentration_1 + drug_concentration_2),
      .by_group = TRUE
    ) %>%
    slice_head(n = 1) %>%
    ungroup()
}

# Define two ways we want to aggregate across concentrations:
# 1. Using the concentration combination with the lowest GR
# 2. Combo with highest log kills
growth_data_bliss_agg_nested <- tribble(
  ~sum_fn, ~col, ~name,
  min, quo(gr), "gr_min",
  max, quo(log_kills), "log_kills_max",
) %>%
  mutate(data = pmap(., calc_summary_stats))

growth_data_bliss_agg <- growth_data_bliss_agg_nested %>%
  select(agg_name = name, data) %>%
  unnest(data) %>%
  select(
    agg_name, panel_id, cell_id, starts_with("lspci_id"),
    starts_with("gr"), starts_with("excess_bliss"),
    starts_with("log_kills"), starts_with("fraction_killing")
  ) %>%
  pivot_longer(
    c(starts_with("gr"), starts_with("excess_bliss"),
    starts_with("log_kills"), starts_with("fraction_killing"), -gr_mean, -excess_bliss_auc)
  ) %>%
  mutate(name = paste(name, agg_name, sep = "_")) %>%
  select(-agg_name) %>%
  pivot_wider(
    names_from = name, values_from = value
  )

library(rasterly)
plot_scatter_rasterly <- function(
  data, x, y, out, out_dir = dir_res, x_range = NULL, y_range = NULL, ...
) {
  args <- rlang::list2(...)
  x_sym <- enquo(x)
  y_sym <- enquo(y)
  df <- data %>%
    # select(!!x_sym, !!y_sym) %>%
    drop_na()
  plt <- df %>%
    rasterly(
      mapping = aes(x = !!x_sym, y = !!y_sym),
      plot_width = args$plot_width %||% 1600, plot_height = args$plot_height %||% 1200,
      x_range = sort(x_range), y_range = sort(y_range)
    ) %>%
    rasterly_points(color = viridisLite::inferno(256)) %>%
    rasterly_build()
  sum_df <- image2data(
    plt$agg$rasterlyPoints1[[1]] %>% {
      .[rev(seq_len(nrow(.))), ]
    },
    x_range = plt$x_range, y_range = plt$y_range, background = 0
  ) %>%
    arrange(color)
  plt <- sum_df %>%
    ggplot(aes(x, y, color = color)) +
    # stat_bin_2d(geom = "raster", bins = c(1200, 1600)) +
    geom_point(shape = 16, size = 0.3, alpha = 0.8) +
    lims(x = x_range, y = y_range) +
    labs(x = args$xlab %||% rlang::as_name(x_sym), y = args$ylab %||% rlang::as_name(y_sym), color = "Count") +
    scale_color_viridis_c(trans = "log10", option = "inferno") +
    theme_minimal() +
    theme(plot.background = element_rect(fill = "white"))
  ggsave(
    file.path(out_dir, paste0(out, ".png")), plt,
    width = 10, height = 6.7
  )
  plt
}

plot_scatter_rasterly(
  growth_data_bliss, gr, excess_bliss, "nci_almanac_gr_vs_excess_bliss_scatter",
  x_range = c(1.5, -3), y_range = c(-1, 1),
  xlab = "GR", ylab = "Excess over Bliss"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, gr_gr_min, excess_bliss_auc, "nci_almanac_gr_min_vs_auc_scatter",
  x_range = c(1.5, -2), y_range = c(-5, 5),
  xlab = "GR min", ylab = "Excess over Bliss AUC"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, gr_mean, excess_bliss_auc, "nci_almanac_gr_mean_vs_auc_scatter",
  x_range = c(1.5, -0.5), y_range = c(-5, 5),
  xlab = "GR mean", ylab = "Excess over Bliss AUC"
)

plot_scatter_rasterly(
  growth_data_bliss_agg, gr_gr_min, excess_bliss_gr_min, "nci_almanac_gr_min_vs_excess_bliss_gr_min",
  x_range = c(1.5, -2), y_range = c(-1, 1),
  xlab = "GR min", ylab = "Excess over Bliss at GR min"
)


plot_scatter_rasterly(
  growth_data_bliss_agg, log_kills_log_kills_max, excess_bliss_log_kills_max, "nci_almanac_log_kills_max_vs_excess_bliss_log_kills_max",
  x_range = c(-0.2, 2), y_range = c(-1, 1),
  xlab = "Max log kills", ylab = "Excess over Bliss at max log kills"
)

## By cell line
dir_res_by_cell <- file.path(dir_res, "by_cell_line")
dir.create(dir_res_by_cell, showWarnings = FALSE)

x <- growth_data_bliss_auc %>%
  filter(
    panel_id == 9, cell_id == 18,
    log_kills_max > 1.5, excess_bliss_log_kills_max > 0
  )

y <- growth_data_bliss %>%
  filter(
    panel_id == 9, cell_id == 18,
    log_kills > 1.5, excess_bliss > 0,
    lspci_id_1 == 55206
  )

scatter_plots_per_cell_line <- growth_data_bliss %>%
  distinct(panel_id, cell_id) %>%
  pmap(
    function(panel_id, cell_id) {
      cell_info <- meta_cell_line %>%
        mutate(across(c(cell_name, panel_name), str_replace_all, r"([^\w])", "_")) %>%
        filter(cell_id == !!cell_id, panel_id == !!panel_id) %>% {
          paste(.$panel_name, .$cell_name, sep = "_")
        }
      
      print(cell_info)
      
      plots <- list()
      
      plots[["gr_vs_excess_bliss"]] <- plot_scatter_rasterly(
        filter(growth_data_bliss, cell_id == !!cell_id, panel_id == !!panel_id),
        gr, excess_bliss, paste0("nci_almanac_", cell_info, "_gr_vs_excess_bliss_scatter"),
        x_range = c(1.5, -3), y_range = c(-1, 1),
        xlab = "GR", ylab = "Excess over Bliss",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      df_auc <- filter(growth_data_bliss_auc, cell_id == !!cell_id, panel_id == !!panel_id)
      plots[["gr_min_vs_auc"]] <- plot_scatter_rasterly(
        df_auc, gr_min, AUC, paste0("nci_almanac_", cell_info, "_gr_min_vs_auc_scatter"),
        x_range = c(1.5, -3), y_range = c(-5, 5),
        xlab = "GR min", ylab = "Excess over Bliss AUC",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["gr_mid_vs_excess_bliss_gr_mid"]] <- plot_scatter_rasterly(
        df_auc, gr_mid, excess_bliss_gr_mid, paste0("nci_almanac_", cell_info, "_gr_mid_vs_excess_bliss_gr_mid_scatter"),
        x_range = c(1.5, -3), y_range = c(-1, 1),
        xlab = "GR mid concentration", ylab = "Excess over Bliss mid concentration",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["gr_mean_vs_auc"]] <- plot_scatter_rasterly(
        df_auc, gr_mean, AUC, paste0("nci_almanac_", cell_info, "_gr_mean_vs_auc_scatter"),
        x_range = c(1.5, -3), y_range = c(-5, 5),
        xlab = "GR mean", ylab = "Excess over Bliss AUC",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["gr_min_vs_auc"]] <- plot_scatter_rasterly(
        df_auc, gr_min, AUC, paste0("nci_almanac_", cell_info, "_gr_min_vs_auc_scatter"),
        x_range = c(1.5, -3), y_range = c(-5, 5),
        xlab = "GR min", ylab = "Excess over Bliss AUC",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
       plots[["gr_min_vs_excess_bliss_gr_min"]] <- plot_scatter_rasterly(
        df_auc, gr_min, excess_bliss_gr_min, paste0("nci_almanac_", cell_info, "_gr_min_vs_excess_bliss_gr_min"),
        x_range = c(1.5, -3), y_range = c(-1, 1),
        xlab = "GR min", ylab = "Excess over Bliss at GR min",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots[["log_kills_max_vs_excess_bliss_log_kills_max"]] <- plot_scatter_rasterly(
        df_auc, log_kills_max, excess_bliss_log_kills_max, paste0("nci_almanac_", cell_info, "_log_kills_max_vs_excess_bliss_log_kills_max"),
        x_range = c(-0.2, 3), y_range = c(-1, 1),
        xlab = "Max log kills", ylab = "Excess over Bliss at max log kills",
        out_dir = dir_res_by_cell, plot_width = 800, plot_height = 600
      )
      
      plots
    }
  )


```

```{r response_rate}
growth_data_bliss_cell_agg <- tibble(response_threshold = c(0.5, 0.95)) %>%
  mutate(
    data = map(
      response_threshold,
      function(response_threshold) {
        growth_data_bliss_agg %>%
          group_by(
            across(c(panel_id, starts_with("lspci_id")))
          ) %>%
          summarize(
            across(
              c(starts_with("excess_bliss"), starts_with("fraction_killing")),
              ~sum(.x > response_threshold) / length(.x)
            ), .groups = "drop"
          )
      }
    )
  ) %>%
  unnest(data)

plt <- growth_data_bliss_cell_agg %>%
  ggplot(
    aes(fraction_killing_gr_min, excess_bliss_gr_min)
  ) +
    geom_point() +
    facet_wrap(~response_threshold) +
    labs(x = "Response rate at GR min", y = "Excess over independence at GR min") +
    # lims(x = c(0, 1), y = c(-1, 1)) +
    theme_minimal()

plot_scatter_rasterly(
  filter(growth_data_bliss_cell_agg, panel_id == 1, response_threshold == 0.95), fraction_killing_gr_min, excess_bliss_gr_min, "nci_almanac_cell_agg_response_rate_gr_min_vs_excess_independence_gr_min",
  x_range = c(0, 1), y_range = c(-1, 1),
  xlab = "Response rate at GR min", ylab = "Excess over independence at GR min"
)

```


```{r proportion_synergistic}
plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(gr_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  ggplot(aes(gr_bin, proportion_synergistic_combos)) +
    geom_col()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(gr_bin) %>%
  summarize(
    proportion_synergistic_combos_per_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  ggplot(aes(gr_bin, proportion_synergistic_combos_per_bin)) +
    geom_col()

plt <- growth_data_bliss %>%
  mutate(
    log_kills_bin = cut(log_kills, c(-Inf, seq(0.5, 9.5, by = 0.5), Inf), ordered_result = TRUE)
  ) %>%
  group_by(log_kills_bin) %>%
  summarize(
    proportion_synergistic_combos_per_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  ggplot(aes(log_kills_bin, proportion_synergistic_combos_per_bin)) +
    geom_col()

plt <- growth_data_bliss_auc %>%
  mutate(
    gr_mean_bin = cut(gr_mean, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_mean_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss_mean > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss_mean > log10(2)) / sum(.$excess_bliss_mean > log10(2)),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_mean_bin, proportion_synergistic_combos_in_bin)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss_auc %>%
  mutate(
    gr_mean_bin = cut(gr_mean, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_mean_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss_mean > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss_mean > log10(2)) / sum(.$excess_bliss_mean > log10(2)),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_mean_bin, proportion_synergistic_combos_in_bin)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    ggbeeswarm::geom_quasirandom(width = 0.2) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_bin) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_bin, proportion_synergistic_combos_in_bin)) +
    # geom_crossbar(stat = "summary", color = "gray40") +
    # ggbeeswarm::geom_quasirandom(width = 0.2) +
    geom_line(aes(group = cell_id)) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_bin) %>%
  filter(n() > 10) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    excess_bliss_mean = mean(excess_bliss),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_bin, excess_bliss_mean)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    # ggbeeswarm::geom_quasirandom(width = 0.2) +
    geom_line(aes(group = cell_id)) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

plt <- growth_data_bliss %>%
  mutate(
    gr_bin = cut(gr, c(-Inf, seq(-3, 1.5, by = 0.5), Inf), ordered_result = TRUE) %>%
      fct_rev()
  ) %>%
  group_by(panel_id, cell_id, gr_bin) %>%
  filter(n() > 10) %>%
  summarize(
    proportion_synergistic_combos_in_bin = sum(excess_bliss > log10(2)) / n(),
    proportion_synergistic_combos = sum(excess_bliss > log10(2)) / sum(.$excess_bliss > log10(2)),
    excess_bliss_mean = mean(excess_bliss),
    .groups = "drop"
  ) %>%
  left_join(
    meta_cell_line
  ) %>%
  ggplot(aes(gr_bin, excess_bliss_mean)) +
    geom_crossbar(stat = "summary", color = "gray40") +
    # ggbeeswarm::geom_quasirandom(width = 0.2) +
    geom_line(aes(group = cell_id)) +
    facet_wrap(vars(panel_name)) +
    theme_minimal()

```



```{r umap}
library(umap)

get_mat <- function(df, var) {
  var_sym <- enquo(var)
  df_long <- df %>%
    transmute(
      panel_cell_id = paste(panel_id, cell_id, sep = "_"),
      lspci_id_combo = paste(lspci_id_1, lspci_id_2, sep = "_"),
      !!var_sym
    )
  df_wide <- df_long %>%
    pivot_wider(names_from = lspci_id_combo, values_from = !!var_sym)
  mat <- df_wide %>%
    column_to_rownames("panel_cell_id") %>%
    as.matrix()
  mat
}

impute_mat <- function(mat, ...) {
  # imp_formula <- reformulate(paste0("`", colnames(mat), "`"))
  # Hmisc::aregImpute(
  #   imp_formula, data = as.data.frame(mat), n.impute = 5, ...
  # )
  VIM::kNN(mat, ...)
}

gr_min_mat <- get_mat(growth_data_bliss_auc, gr_min)
gr_min_mat_imp <- impute_mat(gr_min_mat)
  

umap_gr_min <- umap(
  growth_data_bliss_auc %>%
    transmute(panel_cell_id = paste(panel_id, cell_id, sep = ":"), lspci_id_1, lspci_id_2, gr_min) %>%
    pivot_wider(names_from = panel_cell_id, values_from = gr_min) %>%
    select(-starts_with("lspci_id")) %>%
    as.matrix()
)
```

* Net killing vs Excess over Bliss, using all concentrations
* Or using AUC, adding up all values
* Stratify by cell line
* Use GR value for x-axis
* Range of Excess over Bliss across cell line
* Plot minimum GR vs excess killing at the same doses

* UMAP of AUC values per cell line
